\newpage
\section{Literature}
 \begin{thebibliography}{9}
	\bibitem{budding02}
		Rachel B. Brem, et al.; 2002.
		\emph{Genetic Dissection of Transcriptional Regulation in Budding Yeast}
		Science 296, 752.
	\bibitem{eQTL1}
		Breitling R, Li Y, Tesson BM, Fu J, Wu C, et al.; 2008 
		\emph{Genetical Genomics: Spotlight on QTL Hotspots.}
		PLoS Genetics 4 (10).
	\bibitem{eQTL2}
		Bystrykh L, Weersing E, Dontje B, Sutton S, Pletcher MT, Wiltshire T, Su AI, Vellenga E, Wang J, Manly KF, Lu L, Chesler EJ, Alberts R, Jansen RC, Williams RW, Cooke MP, de Haan G.; 2005
		\emph{Uncovering regulatory pathways that affect hematopoietic stem cell function using genetical genomics.}
		Nature Genetics  37, 225 - 232 
	\bibitem{eQTL3}
		Li, Y.; Alda Alvarez, O.; Gutteling, E.W.; Tijsterman, M.; Fu, J.; Riksen, J.A.G.; Hazendonk, E.; Prins, J.C.P.; Plasterk, R.H.A.; Jansen, R.C.; Breitling, R.; Kammenga, J.E.; 2006
		\emph{Mapping Determinants of Gene Expression Plasticity by Genetical Genomics in C. elegans}
		Plos Genetics 2 (12). - p. 0001 - 0007. 		
	\bibitem{broman09}
		Broman, K.W.; 2009.
		\emph{A brief tour of R/qtl}
		http:$\backslash$$\backslash$www.rQTL.org, R/qtl tutorials.
	\bibitem{broman03}
		Broman, K.W.; Wu, H.; Sen, S.; Churchill, G.A.; 2003.
		\emph{R/qtl: QTL mapping in experimental crosses}
		Bioinformatics, 19:889-890.
	\bibitem{keurtj}
		Keurentijes JJB, Fu J, de Vos CHR,Lommen A, Jansen RC et al; 2006.
		\emph{The genetics of plant metabolism.}
		Nature Genetics. 38, 842–849. 
	\bibitem{Oats}
		L.S. O'Donoughue, S.F. Kianian, P.J. Rayapati, G.A. Penner, M.E. Sorrells,
		S.D. Tanksley, R.L. Phillips, H.W. Rines, M. Lee, G. Fedak, S.J. Molnar, 
		D. Hoffman, C.A. Salas, B. Wu, E. Autrique and A. Van Deynze;  1995.  
		\emph{A molecular linkage map of cultivated oat.}
		Genome 38:368-380
	\bibitem{Hyper}
		Sugiyama, F., Churchill, G. A., Higgens, D. C., Johns, C., Makaritsis, K. P., Gavras, H. Paigen, B; 2001.
		\emph{Concordance of murine quantitative trait loci for salt-induced hypertension with rat and human loci.}
		Genomics 71, 70–77. 
	\bibitem{Listeria}
		Boyartchuk, V. L., Broman, K. W., Mosher, R. E., D'Orazio S. E. F., Starnbach, M. N. Dietrich, W. F; 2001.
		\emph{Multigenic control of Listeria monocytogenes susceptibility in mice.}
		Nature Genetics 27, 259–260. 
	\bibitem{jansen07}
		Jansen R. C.; 2007.
		\emph{Chapter 18 - Quantitative trait loci in inbred lines} 
		Handbook of Stat. Genetics 3th edition,(c) 2007 John Wiley \& Sons, Ltd.
	\bibitem{tierney04}	
		Tierney, L.; Rossini, A.; Li, N.; Sevcikova, H.; 2004.
		\emph{The snow Package: Simple Network of Workstations} Version 0.2-1. 
	\bibitem{tierney03}
		Rossini, A.; Tierney, L.; Li, N.; 2003.
		\emph{Simple parallel statistical computing}
		R. UW Biostatistics working paper series University of Washington. 193
	\bibitem{jansen01}	
		Jansen R. C.; Nap J.P.; 2001
		\emph{Genetical genomics: the added value from segregation}
		Trends in Genetics, 17, 388-391.
	\bibitem{jansen94}		
		Jansen R. C.; Stam P.; 1994
		\emph{High resolution of quantitative traits into multiple loci via interval mapping}
		Genetics, 136, 1447-1455. 
	\bibitem{Churchill94}
		Churchill, G. A.; Doerge, R. W.; 1994
		\emph{Empirical threshold values for quantitative trait mapping}
		Genetics 138, 963–971. 
	\bibitem{jansen93}
		Jansen R. C.; 1993
		\emph{Interval mapping of multiple quantitative trait loci}
		Genetics, 135, 205–211.
	\bibitem{PBS2000}
		Veridian Information Solutions; 2000
		\emph{Portable Batch System - OpenPBS Release 2.3:Administrator Guide}
		website at: www.veridian.com.
	\bibitem{Dempster77}
		Dempster, A. P.; Laird, N. M. and Rubin, D. B.; 1977 
		\emph{Maximum likelihood from incomplete data via the EM algorithm}
		J. Roy. Statist. Soc. B, 39, 1–38.
	\bibitem{Rcurl08}
		Temple Lang D. ; 2008
		\emph{R as a Web Client – the RCurl package}
		Journal of Statistical Software
	\bibitem{brouwer09}
		Brouwer R. 2008/2009;
		\emph{Code from various research projects}
		GBB: MolGen Groningen
	\bibitem{morris07}
		Swertz MA, Jansen R. C.; 2007
		\emph{Beyond standardization: dynamic software infrastructures for systems biology}
		Nat Rev Genet. 3, 235–243. 
	\bibitem{morris04}
		Swertz M.A., De Brock E.O., Van Hijum S.A., De Jong A., Buist G., Baerends R.J., Kok J., Kuipers O.P., Jansen R.C.; 2004.
		\emph{Molecular Genetics Information System (MOLGENIS): Alternatives in developing local experimental genomics databases}
		Bioinformatics,13, 2075–2083.
	\bibitem{mapQTL}
		Van Ooijen \& Maliepaard; 1996
		\emph{MapQTL version 3.0: Software for the calculation of QTL positions on genetic maps.}
		Plant Genome IV Abstracts, http://www.intl-pag.org.
	\bibitem{Rpara}
		Gonzalo V., Jansen R.C.  and Suppi R.L. ; 2008
		\emph{R/parallel – speeding up bioinformatics analysis with R}
		BMC Bioinformatics 2008, 9:390
\end{thebibliography}